---
title: "Gun Control Chi-Square"
author: "Sam Csik"
date: "11/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'm Sam and I'm adding some text here to see if my collaborators can get this file.

It's Sam again, and I'm updating some text.

Ayyyyyyyyyy.

Sean here.

Creator (Sam):

Getting the gc_df.csv data, loading the tidyverse, and getting some summary counts.

```{r}
# load required packages
library(tidyverse)

# load data
gc_df <- read_csv("gc_df.csv")

# QUESTION: Does the sex of the respondent sig effect their perception on ease of gun access; is there a sig. association between sex and perception on guns
  # 2 categorical variables (sex, perception)

# remove NA values & create contingency table
get_guns <- gc_df %>% 
  filter(gender != "NA",
         make_get_guns != "NA") %>% 
  count(gender, make_get_guns) %>% # allows you to set groupings for different categorical variables
  spread(make_get_guns, n) %>% # make into wide format; converts levels within one column so that they now exist in their own columns
  select(-gender) 

rownames(get_guns) <- c("Female", "Male") # change names of table row
```

Collaborator (An): doing chi-square  
- Is there a significant difference in gun control perspectives (as related to access to guns) for males and females?  
- Is there a significant association between sex and gun control recommendations?  
- Is perception of gun control access independent of sex?

```{r chi square proportion table}
# prop.table makes a table of proportions by row
# make data frame matrix in order to work with prop.table()
# calculate proportions across rows with ,1
guns_prop <- prop.table(as.matrix(get_guns), 1)

guns_prop
```

```{r performing chi square}
# use chisq.test on contingency table
guns_x2 <- chisq.test(get_guns)

guns_x2
```
H0: sex is independent of gun control perception.  
HA: sex is non-independent of gun control perception (proportions differ between sexes).

Recommended government action on gun access differs signficantly for males and females ($\chi^2$(`r guns_x2$parameter`) = `r guns_x2$statistic`, p < 0.001.)
















